Optimal Fitting of Planar Curves to Prescribed Constraints

### 2014-04-02 idp/documents/documentation/paper.tex
Included Yannick's feedback.

				\begin{itemize}
				  	\item Platform independence - the project is not bound to any specific platform and can be used on all major operating systems, provided the libraries Kurve depends on are available. 
				  	\item C\# is a high-level language - this reduces overhead in the development process, for example spending time on manual memory management.
					\item Ability to interface with native code - native code can be interfaced with by using reasonable amounts of boilerplate code. As it was clear that native libraries would be used, this is a key requirement.
					\item Familiarity - the developers are familiar with this language. This choice eliminates overhead that would be caused by adopting an unfamiliar language.
					\item UI frameworks - part of the project is a proof of concept user interface to work with curves. Several UI frameworks are available that interface directly with C\#.
				\end{itemize}

@@

				\begin{itemize}
				  	\item Ipopt\footnote{https://projects.coin-or.org/Ipopt} is a library that does numeric non-linear optimization. It allows to solve optimization problems as described in \ref{section:optimization_problem}.
					\item CasADi\footnote{http://casadi.org} provides functionality to perform automatic differentiation of terms. It interfaces with Ipopt directly, which makes it a perfect match for Kurve's symbolic differentiation needs. CasADi is written in C++, but unfortunately only C functions can be called easily from C\#. For that reason, a small CasADi wrapper written in C was required to implement Kurve's C\# term APIs. 
				\end{itemize}

				 Kurve has strong requirements for both libraries: Terms used in the functions that need to be optimized in Kurve are quite large and plentiful, but differentiation and optimization still need to be performed in a short period of time to allow real-time editing of curves. Fortunately, CasADi and Ipopt are very efficient and allow Kurve to offer close to real time editing of curves.

				For Kurve's user interface, GTK\#\footnote{http://www.mono-project.com/GtkSharp} was used. GTK\# is used to draw the curves and controls into a window, as well as handling keyboard and mouse events.

@@

			\subsection{Numerical Optimization}
				The \verb|Curve.Curves| project contains the code required to transform input specifications into curves. It relies on CasADi (through \verb|Krach.Wrappers.Casadi|) to represent curves symbolically and on Ipopt for optimization.

@@

					\begin{itemize}
						\item Any number of point, direction and curvature specifications. These \verb|CurveSpecification|s are stored in a \verb|BasicSpecification|.
						\item The curve's total length is also part of a \verb|BasicSpecification|.
						\item The curves are represented using a number of segments, each of which is represented by some function, realized using the \verb|FunctionTermCurve| class. Because of that, the number of segments as well as the function type that is used to represent each segment are part of \verb|BasicSpecification| as well. 
						The type of function used is represented using a \verb|FunctionTermCurveTemplate| object; this template allows the creation of \verb|FunctionTermCurve| objects.  
						However, this information is only used during optimization; if the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed. It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (i.e. more degrees of freedom mean the curve can satisfy more curve specifications while maintaining constant speed and good fairness, but the optimization process takes longer).
						\item Curves could be generated from a \verb|BasicSpecification| alone. However, if there are multiple curves that fit a \verb|BasicSpecification| very well (i.e. the optimization problem has multiple local optima), different actual curves could be generated from a \verb|BasicSpecification|. This results in a number of problems: e.g., when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process. This is clearly undesirable and a way to disambiguate between the different optima is required. 
						For that reason, a full \verb|Specification| consists of a \verb|BasicSpecification| as well as some disambiguation information. That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded. This both ensures that the same curve will be generated and speeds up the optimization process as well.
					\end{itemize}

@@

					\begin{description}
						\item[Segments]
						In the first step curve segments are created based on the number of segments and the segment template specified in the \verb|BasicSpecification|. A list of parameters of all segments is created; by applying concrete values to each of the parameters, a concrete curve can be obtained. The generated segments are cached and do not have to be regenerated if the number of segments and segment template specified in the basic specification do not change.

						\item[Optimization Problem]
						Next, the optimization problem is created. The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_error}. Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function. This step also computes the first and second derivatives of those functions. The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced. For the total length of the curve a placeholder is used as well. This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed. Thus, the optimization problem is cached and only rebuilt if the optimization segments change or if the number of each type of constraint on the curve changes. If the value of a constraint changes, but the type of constraint remains the same (e.g. a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.

@@

						\item[Solver]
						In this step, a solver for the optimization problem is created. The solver used is provided by Ipopt. A new solver needs to be created whenever either the optimization problem or the substitutions change.
						\item[Position]
							The position represents a result of an optimization process. To obtain this, the solver created in the previous step is run; as the starting position, the previous optimization's position is used. Using the obtained position, a concrete curve can be obtained from the optimization segments. This curve is the desired end result returned by the Optimizer.
					\end{description}

@@

					The \verb|CurveOptimizer| yields this type of curve. The reason why this kind of discrete curve is used is because the \verb|Curve| objects and the \verb|Optimizer| (from the \verb|Kurve.Curves| subsystem) cannot be used at the same time. This, in turn, is caused by the CasADi library, which is not thread safe. However, CasADi is used for representing \verb|Curve|s internally, as well as for performing optimization tasks.

@@

					\item \verb|CurveOptimizer| and \verb|OptimizationWorker| - As the optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), optimization is performed in a background thread. Each \verb|CurveComponent| uses a \verb|CurveOptimizer| to submit changed specifications. A single shared \verb|OptimizationWorker| handles the synchronization and makes sure each submitted specification gets processed one by one. It is necessary to ensure that no optimization processes run in parallel, as the CasADi library (which is used in this process) is not thread safe.

@@

				The user interface does not focus on user friendliness. Its purpose is to test the validity of the approach to curve design researched in this project. For that reason, most actions are performed using keyboard shortcuts. Many user interface elements can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only. It is possible to interact with some components using the mouse. However, there is no toolbar or similar user interface elements that would make any of the features more accessible.

@@

					\begin{itemize}
						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one. ()
						\item To change the degree of the polynomial used in each segment, press alt + `1' to decrease, and alt + `2' to increase the degree by one.
					\end{itemize}
					Keep in mind that the segment count and the degree of the polynomials are not directly visible to the user and are used in the optimization process only. Adding more segments or increasing the degree of the polynomials allows more complicated curves, but has negative effects on the performance.
				\subsubsection{Specification Control}
				\label{section:specification_control}

@@

					A segment also displays the speed of the curve if it differs from the optimal speed. Too high speed is indicated by a red hue; this can usually be fixed by inserting additional length. Too low speed is indicated by a blue hue; usually there will be both blue and red sections. This usually indicates that the curve is too complex (i.e. there are too many specifications) to achieve a good representation with the current curve template. This can be fixed by either removing specifications, adding internal segments to the curve, or increasing the degree of the polynomial function template used in the curve.

>>>>

				\begin{itemize}
				  	\item Platform independence - the project is not bound to any specific platform and can be used on all major operating systems, provided the libraries Kurve depends on are available. 
				  	\item C\# is a high-level language - this reduces overhead in the development process, for example spending time on manual memory management.
					\item Ability to interface with native code - native code can be interfaced with by using reasonable amounts of boilerplate code. Native libraries are used in Kurve, therefore this is a key requirement.
					\item Familiarity - the developers are familiar with this language. This choice eliminates overhead that would be caused by adopting an unfamiliar language.
					\item UI frameworks - part of the project is a proof of concept user interface to work with curves. Several UI frameworks are available that interface directly with C\#.
				\end{itemize}

@@

				\begin{itemize}
				  	\item Ipopt\footnote{https://projects.coin-or.org/Ipopt} is a library that does numeric non-linear optimization. It allows to solve optimization problems as described in \ref{section:optimization_problem}.
					\item CasADi\footnote{http://casadi.org} provides functionality to perform automatic differentiation of terms. It interfaces with Ipopt directly, which makes it the preferred choice for the symbolic differentiation required in Kurve. CasADi is written in C++, but unfortunately only C functions can be called easily from C\#. For that reason, a small CasADi wrapper written in C was required to implement Kurve's C\# term APIs. 
				\end{itemize}

				 Kurve has strong requirements for both libraries: Terms used in the functions that need to be optimized in Kurve are large and plentiful, but differentiation and optimization still need to be performed in a short period of time to allow real-time editing of curves. Fortunately, CasADi and Ipopt are very efficient and allow Kurve to offer close to real time editing of curves.

				For Kurve's user interface, GTK\#\footnote{http://www.mono-project.com/GtkSharp} was used. GTK\# provides functionality to draw the curves and controls into a window, as well as handling keyboard and mouse events.

@@

			\subsection{Numerical Optimization}
				The \verb|Kurve.Curves| project contains the code required to transform input specifications into curves. It relies on CasADi (through \verb|Krach.Wrappers.Casadi|) to represent curves symbolically and on Ipopt for optimization.

@@

					\begin{itemize}
						\item Any number of point, direction and curvature specifications. These \verb|CurveSpecification|s are stored in a \verb|BasicSpecification|.
						\item The curve's total length is also part of a \verb|BasicSpecification|.
						\item The curves are represented using a number of segments, each of which is represented by some function, realized using the \verb|FunctionTermCurve| class. Because of that, the number of segments as well as the function type that is used to represent each segment are part of \verb|BasicSpecification| as well. 
						The type of function used is represented using a \verb|FunctionTermCurveTemplate| object. This template allows the creation of \verb|FunctionTermCurve| objects.  
						However, this information is only used during optimization. If the number of segments and function template yield a curve with enough degrees of freedom with respect to the other specifications, roughly the same curve will be generated even if the number of segments and function template is changed. It is possible to specify the segment count and function template to modify the degrees of freedom that the resulting curves offers (i.e. more degrees of freedom mean the curve can satisfy more curve specifications while maintaining constant speed and good fairness, but the optimization process takes longer).
						\item Curves could be generated from a \verb|BasicSpecification| alone. However, if there are multiple curves that fit a \verb|BasicSpecification| very well (i.e. the optimization problem has multiple local optima), different actual curves could be generated from a \verb|BasicSpecification|. One problem that can arise when saving a curve, if only the \verb|BasicSpecification| were saved, it would be possible that a different curve is displayed when loading the curve, because a different local optimum was reached in the optimization process. This is undesirable and a way to disambiguate between the different optima is required. 
						For that reason, a full \verb|Specification| consists of a \verb|BasicSpecification| as well as some disambiguation information. That disambiguation information is the result of the optimization, which can be used as a starting position for the optimization process when a curve is loaded. This ensures that the same curve will be generated and speeds up the optimization process as well.
					\end{itemize}

@@

					\begin{description}
						\item[Segments]
						In the first step curve segments are created based on the number of segments and the segment template specified in the \verb|BasicSpecification|. A list of parameters of all segments is created. By applying concrete values to each of the parameters, a curve can be obtained. The generated segments are cached and do not have to be regenerated if the number of segments and segment template specified in the \verb|BasicSpecification| do not change.

						\item[Optimization Problem]
						Next, the optimization problem is created. The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_error}. Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the curve's specificationsare realized using a constraint function. This step also computes the first and second derivatives of those functions. The concrete values required by the curve specifications are not used. Instead, placeholder variables are introduced. For the total length of the curve a placeholder is used as well. This allows the reuse of the optimization problem if curve specifications or curve length change. The derivatives do not have to be recomputed. Thus, the optimization problem is cached and only rebuilt if the optimization segments change or if the number of each type of constraint on the curve changes. If the value of a constraint changes, but the type of constraint remains the same (e.g. a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.

@@

						\item[Solver]
						In this step, a solver for the optimization problem is created. The solver used is provided by Ipopt. A new solver needs to be created whenever either the optimization problem or the substitutions change.
						\item[Position]
							The position represents a result of an optimization process. To obtain this, the solver created in the previous step is run. As the starting position, the previous optimization's position is used. Using the obtained position, a concrete curve can be obtained from the optimization segments. This curve is the desired end result returned by the Optimizer.
					\end{description}

@@

					The \verb|CurveOptimizer| yields this type of curve. The reason why this kind of discrete curve is used is due to the lack of the ability to use \verb|Curve| and \verb|Optimizer| (from the \verb|Kurve.Curves| subsystem) at the same time. This, in turn, is caused by the CasADi library, which is not thread safe. However, CasADi is used for representing \verb|Curve|s internally, as well as for performing optimization tasks.

@@

					\item \verb|CurveOptimizer| and \verb|OptimizationWorker| - The optimization process required to generate a curve from specifications takes too much time to perform in the main thread (the user interface would be blocked for too long), therefore optimization is performed in a background thread. Each \verb|CurveComponent| uses a \verb|CurveOptimizer| to submit changed specifications. A single shared \verb|OptimizationWorker| handles the synchronization and makes sure each submitted specification gets processed one by one. It is necessary to ensure that no optimization processes run in parallel, as the CasADi library (which is used in this process) is not thread safe.

@@

				The user interface does not focus on user friendliness. Its purpose is to test the validity of the approach to curve design researched in this project. For that reason, most actions are performed using keyboard shortcuts. Many user interface elements can be selected with the mouse. This enables to use keyboard shortcuts that affect that component only. It is possible to interact with some components using the mouse. However, there is no toolbar or similar user interface elements that would make any of the features more accessible.

@@

					\begin{itemize}
						\item To change the number of segments used, press `1' to decrease and `2' to increase the number of segments by one.
						\item To change the degree of the polynomial used in each segment, press alt + `1' to decrease, and alt + `2' to increase the degree by one.
					\end{itemize}
					Keep in mind that the segment count and the degree of the polynomials are not directly visible to the user and are used in the optimization process only. Adding more segments or increasing the degree of the polynomials allows more complicated curves, but has negative effects on the performance.
				\subsubsection{Specification Control}
				\label{section:specification_control}

@@

					A segment also displays the speed of the curve if it differs from the optimal speed. Too high speed is indicated by a red hue. This can usually be fixed by inserting additional length. Too low speed is indicated by a blue hue. Usually there will be both blue and red sections. This may indicate that the curve is too complex (i.e. there are too many specifications) to achieve a good representation with the current curve template. This can be fixed by either removing specifications, adding internal segments to the curve, or increasing the degree of the polynomial function template used in the curve.
